---
title: "layer_stats"
author: "JLM"
date: "1/3/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(reshape2)
```

```{r}
layerStats <- transform(
  read.csv('layer_stats.csv'),
  max_tile = 2 ^ max_exponent)
```

## State Counts

### 2x2
Nothing all that exciting for the 2x2 board:

```{r}
ggplot(
  subset(layerStats, board_size == 2),
  aes(x = layer_sum, y = states)) +
  geom_line(aes(color = factor(max_depth))) +
  facet_wrap(~ max_tile, ncol = 1)
```

### 3x3 
A bit more interesting for the 3x3:

```{r}
ggplot(
  subset(layerStats, board_size == 3),
  aes(x = layer_sum, y = states)) +
  geom_line(aes(color = factor(max_depth))) +
  facet_wrap(~ max_tile, ncol = 1)
```

Let's look only at the last few:

```{r}
ggplot(
  subset(layerStats, board_size == 3 & max_tile >= 256),
  aes(x = layer_sum, y = states)) +
  geom_line(aes(color = factor(max_depth))) +
  facet_wrap(~ max_tile, ncol = 1)
```

For the game to `max_exponent`, the layer prefix until `max_exponent - 1` is pretty much identical to the same prefix for the game to `max_exponent - 1`.

Let's see what it looks like in log scale.

```{r}
ggplot(
  subset(layerStats, board_size == 3 & max_tile >= 256),
  aes(x = layer_sum, y = states)) +
  geom_line(aes(color = factor(max_depth))) +
  scale_y_log10() +
  facet_wrap(~ max_tile, ncol = 1)
```

The first part looks pretty smooth, and it's not that dissimilar to what we saw in the calcs for the 4x4 board to date.

Let's see what it looks like cumulatively.

```{r}
ggplot(
  transform(
    subset(layerStats, board_size == 3 & max_tile >= 256),
    cumStates = ave(states, paste(max_tile, max_depth), FUN = cumsum)),
  aes(x = layer_sum, y = cumStates)) +
  geom_line(aes(color = factor(max_depth))) +
  facet_wrap(~ max_tile, ncol = 1)
```

The resolver seems to have almost no effect on the first part of the curve, but it does have a larger effect on the latter part of the curve.

```{r}
local({
  d <- aggregate(
    states ~ max_depth,
    subset(layerStats, board_size == 3 & max_tile == 1024),
    sum)
  transform(
    d,
    savings = (max(d$states) - states) / max(d$states))
})
```

So, overall we saved about 10% by resolving to depth 2. If we break that up into before and after the 1024 tile:

```{r}
local({
  d <- aggregate(
    states ~ max_depth,
    subset(layerStats, board_size == 3 & max_tile == 1024 & layer_sum >= 1024),
    sum)
  transform(d, savings = (max(d$states) - states) / max(d$states))
})
```

```{r}
local({
  d <- aggregate(
    states ~ max_depth,
    subset(layerStats, board_size == 3 & max_tile == 1024 & layer_sum < 1024),
    sum)
  transform(d, savings = (max(d$states) - states) / max(d$states))
})
```

So, about 10x as much saved after as before. In graph form:

```{r}
local({
  idvar <- c('board_size', 'max_tile', 'layer_sum')
  d <- subset(
    layerStats,
    select = c(idvar, 'max_depth', 'states')
  )
  r <- reshape(
    d, 
    v.names='states',
    timevar='max_depth',
    idvar=idvar,
    direction='wide')
  r <- transform(
    r,
    relative_states_0 = 1,
    relative_states_1 = states.1 / states.0,
    relative_states_2 = states.2 / states.0,
    states.0 = NULL,
    states.1 = NULL,
    states.2 = NULL,
    states.3 = NULL)
  
  m <- melt(r, measure.vars=c(paste0('relative_states_', 0:2)))
  
  ggplot(
    subset(m, board_size == 3 & max_tile == 1024),
    aes(x = layer_sum, y = value)) +
    geom_line(aes(color = variable))
})
```

But that makes the resolution look much better than it is in later stages --- there are fewer states to start with past 1024.

## Values

Basic value plot for the 3x3 game:

```{r}
local({
  m <- melt(
    subset(
      layerStats,
      board_size == 3 & max_depth == 0 & discount == 0.95 & max_tile >= 128,
      select=c(board_size, max_tile, layer_sum,
               min_value, mean_value, max_value)),
    id.vars = c('board_size', 'max_tile', 'layer_sum'))
  ggplot(m, aes(x = layer_sum, y = value)) +
    geom_line(aes(color = variable)) +
    facet_wrap(~ max_tile, ncol = 1)
})
```

Let's see what it looks like in log space:

```{r}
local({
  m <- melt(
    subset(
      layerStats,
      board_size == 3 & max_depth == 1 & discount == 0.95 & max_tile >= 128,
      select=c(board_size, max_tile, layer_sum,
               min_value, mean_value, max_value)),
    id.vars = c('board_size', 'max_tile', 'layer_sum'))
  ggplot(m, aes(x = layer_sum, y = value)) +
    geom_line(aes(color = variable)) +
    scale_y_log10() +
    facet_wrap(~ max_tile, ncol = 1)
})
```
