---
title: "Tee Log Analysis"
author: "JLM"
date: "2/24/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(stringr)
```

I should probably added these summary stats to the info files... but until then, we can get some interesting stuff out of the logs that I saved with `tee`.

```{r}
extractInfo <- function (file) {
  stampFormat <- '%Y-%m-%d %H:%M:%S +0000'
  stampRx <- '^(\\d+-\\d+-\\d+ \\d+:\\d+:\\d+ \\+\\d+):'
  buildRx <- paste(stampRx, 'build (\\d+)-(\\d+): (\\d+) states')
  reduceRx <- paste(stampRx, 'reduce (\\d+)-(\\d+): ([0-9.]+)MiB \\(([0-9.]+)MiB max\\)$')
  dotsRx <- '^([.]+)'
  
  lines <- readLines(file)
  lineNumbers <- seq_along(lines)
  isBuild <- grepl(buildRx, lines)
  isReduce <- grepl(reduceRx, lines)
  
  build <- data.frame(
    str_match(lines[isBuild], buildRx),
    stringsAsFactors = FALSE)[,2:5]
  names(build) <- c('build_stamp', 'sum', 'max_value', 'num_states')
  build <- transform(
    build,
    build_line = lineNumbers[isBuild],
    build_stamp = as.POSIXct(build_stamp, tz = 'UTC', format = stampFormat),
    sum = as.numeric(sum),
    max_value = as.numeric(max_value),
    num_states = as.numeric(num_states)
  )
  
  reduce <- data.frame(
    str_match(lines[isReduce], reduceRx),
    stringsAsFactors = FALSE)[,2:6]
  names(reduce) <- c('reduce_stamp', 'sum', 'max_value', 'total_mb', 'max_mb')
  reduce <- transform(
    reduce,
    reduce_line = lineNumbers[isReduce],
    reduce_stamp = as.POSIXct(reduce_stamp, tz = 'UTC', format = stampFormat),
    sum = as.numeric(sum),
    max_value = as.numeric(max_value),
    total_mb = as.numeric(total_mb),
    max_mb = as.numeric(max_mb)
  )
  
  merge(build, reduce, all = TRUE)
}
# head(extractInfo('div_rem-20170223.txt'), 20)
# head(extractInfo('part-build-20170221.txt'), 20)
```

```{r}
findDurations <- function (info) {
  buildLines <- subset(
    info, !is.na(build_line), select=c(build_line, build_stamp))
  names(buildLines) <- c('line', 'stamp')
  reduceLines <- subset(
    info, !is.na(reduce_line), select=c(reduce_line, reduce_stamp))
  names(reduceLines) <- c('line', 'stamp')
  allLines <- rbind(buildLines, reduceLines)
  
  findDuration <- function (targetLine) {
    stamp <- subset(allLines, line == targetLine)$stamp[1]
    nextStamp <- subset(allLines, line > targetLine)$stamp[1]
    difftime(nextStamp, stamp, units = 'secs')
  }
  
  transform(
    info,
    build_duration = sapply(build_line, findDuration),
    reduce_duration = sapply(reduce_line, findDuration))
}
# findDurations(head(extractInfo('div_rem-20170223.txt'), 20))
```

```{r}
loadInfo <- function () {
  divRemInfo <- findDurations(extractInfo('tees/div-rem-20170223.txt'))
  divRemInfo$strategy <- 'div-rem'
  batchInfo <- findDurations(extractInfo('tees/part-build-20170221.txt'))
  batchInfo$strategy <- 'batch'
  
  rbind(divRemInfo, batchInfo)
}
info <- loadInfo()
```

At least as parameterised, the div-rem approach takes more disk space:

```{r}
ggplot(info, aes(x = sum)) +
  geom_point(aes(y = total_mb, color = strategy, group = strategy)) +
  facet_grid(max_value ~ .)
```

It generates more, smaller files, as expected:

```{r}
ggplot(info, aes(x = sum)) +
  geom_point(aes(y = max_mb, color = strategy, group = strategy)) +
  facet_grid(max_value ~ .)
```

It does not show any particularly consistent performance advantage, at least on these early layers:

```{r}
ggplot(info, aes(x = sum)) +
  geom_point(aes(y = build_duration, color = strategy, group = strategy)) +
  facet_grid(max_value ~ .)
```

It does spend a bit more time reducing, which is expected.

```{r}
ggplot(info, aes(x = sum)) +
  geom_point(aes(y = reduce_duration, color = strategy, group = strategy)) +
  facet_grid(max_value ~ .)
```

Sanity check: the numbers of states agree.

```{r}
ggplot(info, aes(x = sum)) +
  geom_point(aes(y = num_states, color = strategy, group = strategy)) +
  facet_grid(max_value ~ .)

```
